---
title: "20220829_EDA_R_Markdown_Healthcare_Employee_Attrition"
author: "Paul Jones"
date: "2022-08-29"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load packages, include=FALSE}
library(tidyverse)
library(lubridate)
library(zoo)
library(purrr)
library(skimr)
library(kableExtra)
library(caret)
library(ggpubr)
```

```{r import data}
employee_health_raw <- tibble(read_csv("watson_healthcare_modified.csv"))
```

```{r}
employee_health_clean <- employee_health_raw %>% 
  mutate(
    Age = as.numeric(Age)
    ,Attrition = as.factor(Attrition)
    ,BusinessTravel = as.factor(BusinessTravel)
    ,DailyRate = as.numeric(DailyRate)
    ,Department = as.factor(Department)
    ,DistanceFromHome = as.numeric(DistanceFromHome)
    ,Education = as.numeric(Education)
    ,EducationField = as.factor(EducationField)
    ,EmployeeCount = as.numeric(EmployeeCount)
    ,EnvironmentSatisfaction = as.numeric(EnvironmentSatisfaction)
    ,Gender = as.factor(Gender)
    ,HourlyRate = as.numeric(HourlyRate)
    ,JobInvolvement = as.numeric(JobInvolvement)
    ,JobLevel = as.numeric(JobLevel)
    ,JobRole = as.factor(JobRole)
    ,JobSatisfaction = as.numeric(JobSatisfaction)
    ,MaritalStatus = as.factor(MaritalStatus)
    ,MonthlyIncome = as.numeric(MonthlyIncome)
    ,MonthlyRate = as.numeric(MonthlyRate)
    ,NumCompaniesWorked = as.numeric(NumCompaniesWorked)
    ,Over18 = as.factor(Over18)
    ,OverTime = as.factor(OverTime)
    ,PercentSalaryHike = as.numeric(PercentSalaryHike)
    ,PerformanceRating = as.numeric(PerformanceRating)
    ,RelationshipSatisfaction = as.numeric(RelationshipSatisfaction)
    ,StandardHours = as.numeric(StandardHours)
    ,Shift = as.numeric(Shift)
    ,TotalWorkingYears = as.numeric(TotalWorkingYears)
    ,TrainingTimesLastYear = as.numeric(TrainingTimesLastYear)
    ,WorkLifeBalance = as.numeric(WorkLifeBalance)
    ,YearsAtCompany = as.numeric(YearsAtCompany)
    ,YearsInCurrentRole = as.numeric(YearsInCurrentRole)
    ,YearsSinceLastPromotion = as.numeric(YearsSinceLastPromotion)
    ,YearsWithCurrManager = as.numeric(YearsWithCurrManager)
)
```

```{r split data into training and test sets}
a <- createDataPartition(employee_health_clean$Attrition, p = 0.7, list=FALSE)
training <- employee_health_clean[a,]
test <- employee_health_clean[-a,]
```

```{r}
skimr::skim(training)
```

```{r}
str(training)
```

```{r}
# Function to plot the plots
plothist <- function(df,x){
  ggplot(data = df) +
    geom_histogram(aes(x=.data[[x]],fill=Attrition), stat = "count")
}

plotline <- function(df,x){
  ggplot(data = df,aes(x=.data[[x]],group=Attrition,color=Attrition)) +
    geom_line(stat = "count")
}

plotbox <- function(df,x){
  ggplot(data = df,aes(x=.data[[x]],group=Attrition,color=Attrition)) +
    geom_boxplot()
}

plotbar_percent <- function(df,x){
  ggplot(data = df, aes(x = .data[[x]], fill = Attrition)) +
    geom_bar(position = "fill") + 
    ylab("proportion") +
    stat_count(geom = "text", 
               aes(label = stat(count)),
               position=position_fill(vjust=0.5), colour="white") +
    theme(legend.position="bottom", legend.direction = "horizontal")
}

nested_data_type <- function(df){
  df[[1]] %>% 
    is.numeric()
}

ggarrange_fct <- function(plot1,plot2,plot3,plot4){
  ggarrange(plot1, plot2, plot3, plot4, nrow = 2, ncol = 2, common.legend = TRUE, legend="bottom")
}

# Nest data frames for each variable and create plots against attrition
training_nested <- training %>%
  names() %>% 
  tibble(data = .) %>% 
  mutate(nested_df = map(data,~training %>% select(.x,Attrition))) %>% 
  mutate(
    nested_hist = map2(.x=nested_df,.y=data,plothist)
    ,nested_line = map2(.x=nested_df,.y=data,plotline)
    ,nested_box = map2(.x=nested_df,.y=data,plotbox)
    ,nested_bar_per = map2(.x=nested_df,.y=data,plotbar_percent)
    ,data_type_fct = unlist(map(.x=nested_df,nested_data_type))
    ,ggarr = case_when(
      data_type_fct == TRUE ~ pmap(list(plot1 = nested_hist,  plot2 = nested_bar_per, plot3 = nested_box, plot4 = nested_line), ggarrange_fct)
      ,data_type_fct == FALSE ~ map2(.x = nested_hist, .y = nested_bar_per, 
        ~ ggarrange(.x,.y,nrow = 2, common.legend = TRUE, legend="bottom"))
  ))

  rowwise() %>% 
  mutate(
    arrange_plot = list(ggarrange(
      nested_hist
      ,nested_bar_per
      #,print(nested_box)
      #,print(nested_line)
      #,ncol=2
      ,nrow=2
      ,common.legend = TRUE
      ,legend="bottom")
  ))

  
  unnest(cols = c(nested_df)) %>% 
  select(1) %>% 
  is.numeric()
  ggplot(aes(x = BusinessTravel, fill = Attrition)) +
    geom_bar(position = "fill") + 
    ylab("proportion") +
    stat_count(geom = "text", 
               aes(label = stat(count)),
               position=position_fill(vjust=0.5), colour="white")
  
training_nested %>% 
  filter(data == "Age") %>% 
  select(arrange_plot) %>% 
  unnest(cols = c(arrange_plot))

print(training_nested$ggarr[[2]])
```

## Age

```{r}
print(training_nested$nested_hist[[2]])
print(training_nested$nested_box[[2]])
print(training_nested$nested_line[[2]])
print(training_nested$nested_bar_per[[2]])

ggarrange(
  print(training_nested$nested_hist[[2]])
  ,print(training_nested$nested_bar_per[[2]])
  ,print(training_nested$nested_box[[2]])
  ,print(training_nested$nested_line[[2]])
  ,ncol=2
  ,nrow=2
  ,common.legend = TRUE
  ,legend="bottom")


```

```{r}
print(training_nested$nested_box[[2]])
```

```{r}
print(training_nested$nested_line[[2]])
```

```{r}
print(training_nested$nested_bar_per[[2]])
```

## Business Travel

```{r}
print(training_nested$nested_hist[[4]])
```

```{r}
print(training_nested$nested_bar_per[[4]])
```

## Daily Rate

```{r}
print(training_nested$nested_hist[[5]])
```

```{r}
print(training_nested$nested_box[[5]])
```

## Department

```{r}
print(training_nested$nested_hist[[6]])
```

```{r}
print(training_nested$nested_bar_per[[6]])
```


## Distance From Home

```{r}
print(training_nested$nested_hist[[7]])
```

```{r}
print(training_nested$nested_bar_per[[7]])
```

```{r}
print(training_nested$nested_box[[7]])
```

```{r}
print(training_nested$nested_line[[7]])
```

## Education

```{r}
print(training_nested$nested_hist[[8]])
```

```{r}
print(training_nested$nested_bar_per[8])
```

## Education Field

```{r}
print(training_nested$nested_hist[[9]])
```

```{r}
print(training_nested$nested_bar_per[[9]])
```

## EmployeeCount

```{r}
print(training_nested$nested_hist[[10]])
```

Feature adds no additional information to the model.

## Environment Satisfaction

```{r}
print(training_nested$nested_hist[[11]])
```

```{r}
print(training_nested$nested_bar_per[[11]])
```

## Gender

```{r}
print(training_nested$nested_hist[[12]])
```

```{r}
print(training_nested$nested_bar_per[[12]])
```

## Hourly Rate

```{r}
print(training_nested$nested_hist[[13]])
```

```{r}
print(training_nested$nested_bar_per[[13]])
```

```{r}
print(training_nested$nested_box[[13]])
```

```{r}
print(training_nested$nested_line[[13]])
```

## Job Involvement

```{r}
print(training_nested$nested_hist[[14]])
```

```{r}
print(training_nested$nested_bar_per[[14]])
```

## Job Level

```{r}
print(training_nested$nested_hist[[15]])
```

```{r}
print(training_nested$nested_bar_per[[15]])
```

## Job Role

```{r}
print(training_nested$nested_hist[[16]])
```

```{r}
print(training_nested$nested_bar_per[[16]])
```

## Job Satisfaction

```{r}
print(training_nested$nested_hist[[17]])
```

```{r}
print(training_nested$nested_bar_per[[17]])
```

## Marital Status

```{r}
print(training_nested$nested_hist[[18]])
```

```{r}
print(training_nested$nested_bar_per[[18]])
```

## Monthly Income

```{r}
print(training_nested$nested_hist[[19]])
```

```{r}
print(training_nested$nested_box[[19]])
```

## Monthly Rate

```{r}
print(training_nested$nested_hist[[20]])
```

```{r}
print(training_nested$nested_box[[20]])
```

## Num Companies Worked

```{r}
print(training_nested$nested_hist[[21]])
```

```{r}
print(training_nested$nested_bar_per[[21]])
```

```{r}
print(training_nested$nested_line[[21]])
```

```{r}
print(training_nested$nested_box[[21]])
```

## Over 18

```{r}
print(training_nested$nested_hist[[22]])
```

Feature doesn't add any additional information to model.

## Over Time

```{r}
print(training_nested$nested_hist[[23]])
```

```{r}
print(training_nested$nested_bar_per[[23]])
```

## Percent SalaryHike

```{r}
print(training_nested$nested_hist[[24]])
```

## Performance Rating

```{r}
print(training_nested$nested_hist[[25]])
```

## Relationship Satisfaction

```{r}
print(training_nested$nested_hist[[26]])
```

## Standard Hours

```{r}
print(training_nested$nested_hist[[27]])
```

## Shift

```{r}
print(training_nested$nested_hist[[28]])
```

## Total Working Years

```{r}
print(training_nested$nested_hist[[29]])
```

## Training Times Last Year

```{r}
print(training_nested$nested_hist[[30]])
```

## Work Life Balance

```{r}
print(training_nested$nested_hist[[31]])
```

```{r}
print(training_nested$nested_bar_per[[35]])
```

```{r}
print(training_nested$nested_box[[35]])
```

```{r}
print(training_nested$nested_line[[35]])
```

## Years At Company

```{r}
print(training_nested$nested_hist[[32]])
```

```{r}
print(training_nested$nested_bar_per[[32]])
```

```{r}
print(training_nested$nested_box[[32]])
```

```{r}
print(training_nested$nested_line[[32]])
```

## Years In Current Role

```{r}
print(training_nested$nested_hist[[33]])
```

```{r}
print(training_nested$nested_bar_per[[33]])
```

```{r}
print(training_nested$nested_box[[33]])
```

```{r}
print(training_nested$nested_line[[33]])
```

## Years Since Last Promotion

```{r}
print(training_nested$nested_hist[[34]])
```

```{r}
print(training_nested$nested_bar_per[[34]])
```

```{r}
print(training_nested$nested_box[[34]])
```

```{r}
print(training_nested$nested_line[[34]])
```

## Years With Current Manager

```{r}
print(training_nested$nested_hist[[35]])
```

```{r}
print(training_nested$nested_bar_per[[35]])
```

```{r}
print(training_nested$nested_box[[35]])
```

```{r}
print(training_nested$nested_line[[35]])
```

## Modelling

```{r}
rf_model <- randomForest::randomForest(Attrition ~., training[2:35], ntree=50)

rf_model

test %>% 
  mutate(results = predict(rf_model, newdata = test, type = "prob"))

new <- test

bind_cols(test, as_tibble(predict(rf_model, newdata = test, type = "prob")))
as_tibble(predict(rf_model, newdata = test, type = "prob"))



# Predicting the Test set results
y_pred <- predict(rf_model, newdata = test, type = "prob")

# Confusion Matrix
confusion_mtx <- confusionMatrix(y_pred, reference = test$Attrition)
confusion_mtx

# Plotting model
plot(rf_model)

# Importance plot
randomForest::varImpPlot(rf_model)

# Variable importance plot
randomForest::importance(rf_model)

```
